A Study of the Effectiveness of Usage Examples in REST API Documentation

### 2017-07-20 rebuttal.txt

We thank the reviewers for their time and constructive feedback on our paper. We plan to incorporate the suggestions and clarify the concerns as follows:

REVIEW 1

We'd like to clarify that the paper studied 26 developers in two groups. We'll update the title on Table 1 to further clarify this number.

1.1 We agree with the reviewer that a third experiment with an enhanced documentation without adding usage examples may provide different insights. We will mention this as a limitation of this paper in Section V. (Threats to Validity).

1.2 The examples show a single API call for three API actions with their associated responses. The data for the examples are taken from the original API documentation and the unit test code of the API. The original API documentation is located at http://v2.wp-api.org/reference/posts/ and the enhanced documentation is located at https://spyrest.github.io/docs-v2/reference/posts/

1.3 On task T4, average trial API call and time spent by G2 was higher than G1 primarily due to P2.7 who made 11 API calls. The rest of G2 participants had the following numbers of trial API calls (1, 5, 5, 3, 2, 6, 1, 1, 3) on T4. P2.7 reached the correct answer on the third trial, but made 8 subsequent trials to confirm the results.

1.4 In the paper, we performed a qualitative study to understand the types of obstacles faced by REST API client developers without examples with the aim that our results will help the design of API documentation. The findings differ from local APIs as follows: Local APIs are often used by API client developers in the same programming language of the API. As such, information regarding data types and formats can often be inferred using code-completion features. Also, HTTP request and response headers are unique to web APIs that are not used in a local API. We'll clarify these distinctions related to our findings in an the camera ready version.

REVIEW 2

2.1 As we discused in 1.1, we will acknowledge this as a limiation of our work. We have provided links to the original and enhanced API documentation in 1.2. The data presented in the usage examples in the enhanced version is sourced from the original API documentation. For example, in the "Create a Post" section, the original documentation mentions five possible values for the attribute "status". In the enhanced version, we used one of those five values.

2.2 Based on the feedback about quantitative analysis, we performed statistical tests on the four quantiative results and found the following: a) Using Fisher test, we observed a p-value of 0.0003 for the success rates of the two groups, b) p-value of 0.017 for the trial API calls made by the two groups using Wilcoxon rank sum test, c) p-value of 0.003 for the satiscation rating of the two groups using Wilcoxon rank sum test, and d) p-value of 0.051 for the time taken by the two groups using Wilcoxon rank sum test. Based on these results, we see that the observations reported on Figures 4a and 5 are satistically significant, while Figure 4b is not. We will add these findings in the camera-ready version.

2.3 In our study, for three API actions, we added one example API call per action with its request and response data showing the HTTP query parameters, headers and body attributes. We observed these single examples helped developers in performing the API tasks avoiding the obstacles that are faced without the examples. For complicated API actions, it may be necessary to provide multiple examples for a single API action to describe intended usages scenarios. We will mention this as a limitation in our paper.

2.4 We'll update the references in the related work section as suggested and make the grammatical corrections as pointed out by the reviewer.

REVIEW 3

3.1 Please see 1.3

3.2 The participants accessed the web-based REST API explorer used for this study on their own computers and we did not record web search activities. They were told to access the internet search tools as needed only to seek information regarding the given API tasks during their participation.

3.3 We will update the paper to incorporate the suggested corrections in a new version.

----------------------- REVIEW 1 ---------------------
PAPER: 20
TITLE: A Study of the Effectiveness of Usage Examples in REST API Documentation
AUTHORS: S. M. Sohan, Frank Maurer, Craig Anslow and Martin Robillard

Overall evaluation: 2 (accept)
Reviewer's confidence: 4 ((high))

----------- Review -----------
Summary:
This paper evaluates the challenges faced by developers using REST APIs and provides a list of empirically-derived recommendations for addressing those challenges. The paper presents a between-subject study of 16 developers where one treatment got the standard WordPress REST API v2 and the other received an enhanced edition that provided 3 additional usage examples that contained useful additional information not found in the normal API. Participants were given 6 REST API tasks to complete, a together performed a total of 539 API calls. The results showed that participants who received the treatment with additional examples finished more of the tasks than those who didn’t

Review:
This paper was straightforward and well-written. It positioned itself well among the related work and made a compelling argument for how it differed to what’s been done before.

I found the takeaways to be useful, yet I wanted to see more depth from the recommendations beyond “API’s should (1) define expected data types, (2) define data formats, (3) define expected inputs and (4) define prerequisites.” The authors present these takeaways as justifications for including examples in the APIs, but they could also be addressed by having through and complete documentation. It would have been insightful to have the study include a third treatment, one where the existing documentation has the missing information added sans example. I suspect the treatment with examples would still outperform and it would make for a more compelling paper.

I’d like to see all three examples that were provided to participants. It’s unclear in the paper if the three examples were too similar to the six tasks given to participants and thus effectively giving the answers to the participants. I don’t think this is the case, but it would be to verify.

Although the authors already lightly discuss it, I would like to know more about Task 4 where G2 participants spent more time on average than G1 participants. Was it because G1 participants were more likely to not finish that task, whereas G2 participants finished it?

I think one of the difficulties of facing this paper is that it all seems kind of obvious. Add examples to address the questions that participants are likely to have and unsurprisingly, participants do better. This obviousness could have been mitigated by providing some challenges faced by programmers that are unique to REST APIs or implications for design for this unique domain. As far as I can tell, the suggestions provided seem to be similar to the ones in the existing literature for local APIs. However, these are not sufficient reasons to reject the paper, so I would argue for accepting it.

----------------------- REVIEW 2 ---------------------
PAPER: 20
TITLE: A Study of the Effectiveness of Usage Examples in REST API Documentation
AUTHORS: S. M. Sohan, Frank Maurer, Craig Anslow and Martin Robillard

Overall evaluation: -1 (weak reject)
Reviewer's confidence: 5 ((expert))

----------- Review -----------
This paper describes a controlled experiment with 26 experienced
developers using a special-purpose web editor to create REST
calls. Some had usage examples as part of the documentation and some
did not. The people who did not did worse. It is good they they got
"26 participants from sixteen different companies and six different
countries".

It has been known for a long time that examples are helpful and are
preferred by developers, so that is not a new contribution. And the
control condition in this study was clearly-inadequate documentation,
so the fact that these people did worse is no surprise and not
particularly useful to know. It would have been much more valid if the
control condition had well-done, complete documentation. Many, many
previous papers have identified inadequate documentation as a frequent
complaint about APIs, so showing that the official documentation for
these REST APIs is incomplete is not interesting either. I didn't find
any of the Recommendations to be new or non-obvious.

Since you spend a lot of time comparing the two conditions (e.g.,
Figures 4,5), you need to do the statistical analysis to see if there
are actually any statistically significant differences. Any
quantitative comparison is not valid without that.

With respect to Recommdation 4: how many examples are necessary? It
seems like you are requesting that every combination of calls would
need an example. What TYPE of APIs examples are needed?

In related work section 4, you mention: "but it was challenging for
the participants to locate relevant examples" -- did you address this
in your study? If not, no need to mention it. Also, the study in [18]
doesn't seem relevant.

The relate work should include this study, even though it is about
SOAP APIs instead of REST:

Brad A. Myers, Sae Young Jeong, Yingyu Xie, Jack Beaton, Jeff Stylos,
Ralf Ehret, Jan Karstens, Arkin Efeoglu, Daniela K. Busse. "Studying
the Documentation of an API for Enterprise Service-Oriented
Architecture". The Journal of Organizational and End User Computing,
JOEUC, IGI Publishing. vol. 22, no. 1, Jan-Mar, 2010. pp. 23-51. IGI global,
DOI: 10.4018/joeuc.2010101903.

Low level writing issues:

This is a better motivation than what is put in the abstract: "API
developers need to know what to include in the examples and why" - but
you don't actually follow through on doing this.

In R4 on page 2, you say: "with no prior experince of using the
WordPress REST API V2" but you haven't said that is what you are using
yet. Maybe just say "the selected API" here, or better yet - introduce
the target API in the introduction.

typo: documentation of the the

missing word: are required to the same

typo: APIs and and a

Figure 3 - screen shot is too small - can't read.

6) Data Analysis: this section has too much detail, and the analysis
was done wrong anyway (see above).

page 5: "each other to provide consistency" -- better to say "to
evaluate consistency"

Figure 4a - I can't tell which vertical axis goes with which dataset -
are the bars or lines the "success rate"?

----------------------- REVIEW 3 ---------------------
PAPER: 20
TITLE: A Study of the Effectiveness of Usage Examples in REST API Documentation
AUTHORS: S. M. Sohan, Frank Maurer, Craig Anslow and Martin Robillard

Overall evaluation: 2 (accept)
Reviewer's confidence: 3 ((medium))

----------- Review -----------
In this paper, the authors present results from a controlled study they carried out to evaluate the impact of providing usage examples in REST API documentation. While generating and maintaining API documentation together with examples of usage can be expensive, it helps client developers work more efficiently while making fewer mistakes. The authors summarize their findings through:
1) A list of common obstacles client developers who use REST APIs encounter in the absence of usage examples.
2) A list of guidelines on incorporating usage examples in API documentation.
3) Empirical evidence that shows positive impact of usage examples on time taken and success rate for task completion, and overall developer satisfaction.

On page 6, you state "For T4, where the participants...rate." Any insights into why G2 executed more API calls for T4? Given the design of your custom REST API explorer shown in Figure 2, with the response header so prominently displayed, I am curious as to why participants in G2 had to execute more requests to be able to complete the task.

Since participants were allowed to access the internet, did you log those accesses? Any insights into the sort of information G1 and G2 sought through their web searches?

Minor corrections:
- Pg 1, col 2: Change "...obstacles API client..." to "...obstacles do API client..."
- Pg 2, col 1: Change "...images, tags." to "...images, and tags."
- Pg 3, col 1: Change "...requires to the..." to "...required to use the..."
- Pg 7, col 1: Change "Eight the..." to "Eight of the..."

### 2017-07-20 controlled_study.tex
adds all the required changes

\begin{itemize}
  \item \textbf{RQ}. What obstacles API client developers face when using a REST API documentation that lacks usage examples?
\end{itemize}

@@

We selected the WordPress REST API V2 for this study. WordPress is a blog-like open-source (R2) framework used by over 409 million people to visit 23.6 billion pages each month \footnote{\url{https://wordpress.com/activity/}}. The API allows programmatic access to list, create, update, and delete WordPress data such as blog posts, comments, users, images, tags.

@@

To verify the response of an API call, the API client developers can use HTTP response headers and/or HTTP response body. To perform the tasks, the participants are required to use one or more of the inputs I1-4. In the following paragraphs, we describe each task with its description and the study observation goals against the aforementioned API input and output information. For each of the tasks, the participants are required to the same WordPress REST API and one of the two variants of accompanying documentation.

@@

\subsubsection{Data Analysis}
We analyzed the data artifact qualitatively to understand the obstacles that REST API client developers face when using an API documentation that lacks usage examples. We exported a table of data from the web-based REST API explorer. For each row, the table contains the following columns: API task, participant identifier, timestamp, API trial number, I1-4, response headers and body. The values for I1-4 used by the participants were manually coded to group the participant responses into categories. The categories help us determine the information type that API client developers need in the API documentation to perform API tasks. The analysis started with an empty set of codes and new codes were introduced to describe scenarios that did not fall under existing codes. The first author applied the codes on the raw data artifact and provided a coding scheme comprising 11 codes to a co-author. The co-author applied the codes to 82/539 API calls (10\% confidence interval with 95\% confidence level). The two authors validated the codes against each other to provide consistency. The initial coding from the co-author had discrepancies in 4 / 11 codes for 27 / 82 API call samples. After discussing the coding scheme, the discrepancies were resolved as the definition of the terms became clearer.

@@

Task T5 also required the participants to use an the ISO8601 formatted date for publishing a post with a specific date. Even though the original documentation mentions ISO8601 format, it does not provide an example. Instead of using \lstinline{2016-01-01T00:00:00-00:00}, participants from G1 used date formats as follows: \lstinline{January 1, 2016, 2016-01-01T12:00:00, 20160101}. P1.8 mentioned the following: \textit{``Got hung up trying to figure out the formatting of the date (which doesn't appear to be ISO8601, despite what it says)''}. After analyzing the response, we found that P1.8 used the ISO8601 formatted date in the API calls, but did not provide the required \lstinline{second} portion of the time. Eight the 16 participants in G1 had an error in the date formats. With an example, 1 of the 10 participants had the same error from G2.

@@

\subsubsection{Controlled Studies}
Nasehi et al. performed a controlled study to understand if API unit tests can be used to provide as usage examples to facilitate API client developers \cite{5609553}. The researchers found that the examples from the unit tests helped understanding the API concepts better but it was challenging for the participants to locate relevant examples. Endrikat et al. performed a controlled study with four groups of participants to understand the impact of API documentation on APIs that are implemented using programming languages with static and dynamic type systems \cite{Endrikat:2014:ADS:2568225.2568299}. Participants were given a set of failing unit tests and were asked to make them pass by writing code using the studied API. They found that the participant group using a statically typed API with explicit documentation could get the tests to pass in less time than the other groups. Ellis et al. performed a controlled study with two groups of developers to understand the impact of using a constructor vs. a factory method design pattern on API client development time \cite{Ellis:2007:FPA:1248820.1248863}. They found that API client developers needed more time to complete API tasks using factory method compared to using a constructor. Our work in this paper is based on a controlled study and shares part of the setup that were used by the aforementioned controlled studies with the following differences: our goal is to understand the obstacles faced by API client developers without usage examples;  we focus on REST APIs instead of local APIs; and, the participants are all professional software engineers.

>>>>

\begin{itemize}
  \item \textbf{RQ}. What obstacles do API client developers face when using a REST API documentation that lacks usage examples?
\end{itemize}

@@

We selected the WordPress REST API V2 for this study. WordPress is a blog-like open-source (R2) framework used by over 409 million people to visit 23.6 billion pages each month \footnote{\url{https://wordpress.com/activity/}}. The API allows programmatic access to list, create, update, and delete WordPress data such as blog posts, comments, users, images, and tags.

@@

To verify the response of an API call, the API client developers can use HTTP response headers and/or HTTP response body. To perform the tasks, the participants are required to use one or more of the inputs I1-4. In the following paragraphs, we describe each task with its description and the study observation goals against the aforementioned API input and output information. For each of the tasks, the participants are required to use the same WordPress REST API and one of the two variants of accompanying documentation.

@@

\subsubsection{Data Analysis}
We analyzed the data artifact qualitatively to understand the obstacles that REST API client developers face when using an API documentation that lacks usage examples. We exported a table of data from the web-based REST API explorer. For each row, the table contains the following columns: API task, participant identifier, timestamp, API trial number, I1-4, response headers and body. The values for I1-4 used by the participants were manually coded to group the participant responses into categories. The categories help us determine the information type that API client developers need in the API documentation to perform API tasks. The analysis started with an empty set of codes and new codes were introduced to describe scenarios that did not fall under existing codes. The first author applied the codes on the raw data artifact and provided a coding scheme comprising 11 codes to a co-author. The co-author applied the codes to 82/539 API calls (10\% confidence interval with 95\% confidence level). The two authors validated the codes against each other to evaluate consistency. The initial coding from the co-author had discrepancies in 4 / 11 codes for 27 / 82 API call samples. After discussing the coding scheme, the discrepancies were resolved as the definition of the terms became clearer.

@@

Task T5 also required the participants to use an the ISO8601 formatted date for publishing a post with a specific date. Even though the original documentation mentions ISO8601 format, it does not provide an example. Instead of using \lstinline{2016-01-01T00:00:00-00:00}, participants from G1 used date formats as follows: \lstinline{January 1, 2016, 2016-01-01T12:00:00, 20160101}. P1.8 mentioned the following: \textit{``Got hung up trying to figure out the formatting of the date (which doesn't appear to be ISO8601, despite what it says)''}. After analyzing the response, we found that P1.8 used the ISO8601 formatted date in the API calls, but did not provide the required \lstinline{second} portion of the time. Eight of the 16 participants in G1 had an error in the date formats. With an example, 1 of the 10 participants had the same error from G2.

@@

\subsubsection{Controlled Studies}
Nasehi et al. performed a controlled study to understand if API unit tests can be used to provide as usage examples to facilitate API client developers \cite{5609553}. The researchers found that the examples from the unit tests helped understanding the API concepts better but it was challenging for the participants to locate relevant examples. Endrikat et al. performed a controlled study with four groups of participants to understand the impact of API documentation on APIs that are implemented using programming languages with static and dynamic type systems \cite{Endrikat:2014:ADS:2568225.2568299}. Participants were given a set of failing unit tests and were asked to make them pass by writing code using the studied API. They found that the participant group using a statically typed API with explicit documentation could get the tests to pass in less time than the other groups. Our work in this paper is based on a controlled study and shares part of the setup that were used by the aforementioned controlled studies with the following differences: our goal is to understand the obstacles faced by API client developers without usage examples;  we focus on REST APIs instead of local APIs; and, the participants are all professional software engineers.
